\LARGE \bf Energy-Aware Dynamic Mission Planning Algorithm for UAVs

### 2020-10-25 icra-2021.tex
proofreading

\begin{figure*}[t]
  \centering
  \input{\figpath/icra-2021_overview2.tikz}
  \caption{The mission consists of circles and lines that cover a polygon. The UAV heads to $\varphi_0$ initially, using the desired velocity vector $\dot{\mathbf{p}}_d$. It later heads to $\varphi_0^-$ reducing the radius $c_{0,1}$ to satisfy the battery constraints (indeed the energy coefficients $\mathbf{q}_{k_2}$ evolution is higher then $\mathbf{q}_{k_3}$). The UAV then converges to $\varphi_1$ in stage $\mathcal{M}_1$, $\varphi_2$ in stage $\mathcal{M}_2$, and so on (the circle $\varphi_2$ is not visible in the figure; it connects $\varphi_1$ and $\varphi_3$).}
  \label{fig:overview}
\end{figure*}

@@

In the spirit of reducing costs, and resources, we showcase the algorithm using the problem of dynamic mission planning for a precision agriculture fixed-wing UAV. Such a scenario is often put into practice~\cite{hajjaj2014review} with ground mobile robots used for harvesting~\cite{qingchun2012study,dong2011development, de2011design, aljanobi2010setup, li2008analysis, edan2000robotic}, and UAVs for preventing damage and ensuring better crop quality~\cite{puri2017agriculture, daponte2019review}. The mission consists of a UAV flying in circles and lines covering a polygon, detecting obstacles using a convolutional neural network (CNN), and informing grounded mobile robots employed for future harvesting--a survey mission optimized for the craft's dynamics. The algorithm plans the mission controlling the processing rate and the radius of the circles, affecting the distance between the lines. Figure~\ref{fig:overview} shows an initial slice of such a mission. The UAV first heads to a circular TEE with a given radius, which is later reduced, as e.g., windy weather required the adjustment of the control input to avoid battery depletion. Data indicates a potential extension of up to 13 minutes over an hour by merely switching to the lowest computations.

@@

  The \emph{mission} is the union of all the stages, along with a function $\lambda:\mathbb{R}^2\rightarrow\mathcal{M}_i$ which maps a point $\mathbf{p}\in\mathbb{R}^2$ of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$ to a specific stage. For simplicity the system is sampled discrete-time and $l$ is the total number of stages
  \begin{equation}
    \mathcal{M}:=(\lambda(\mathbf{p}),\bigcup_{i\in[l]}{\mathcal{M}_i}),
  \end{equation}
  the algorithm inputs $\mathcal{M}$, initial position, and energy coefficients guess, and outputs the position, the instantaneous energy consumption, and the control input--an action performed evolving the mission state.
\end{defn}

@@

The set
\begin{equation}\label{eq:area}
  \mathcal{P}_i:=\{\mathbf{p}_k\mid\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i\},
\end{equation}
delimits the area where the $i$-th TEE $\varphi_i$ is free to evolve using $\rho$ adjustments $\mathbf{c}_i:=c_{i,1},...,c_{i,\rho}$ (gray area in Figure~\ref{fig:overview}), being the TEE satisfied for all the approaching points $\varphi_i\rightarrow\mathbb{C}_i$.

The algorithm uses the concept to select the adjustments s.t. $\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$ has the highest energy value. It guides the UAV to the new position $\mathbf{p}_{k+1}$ using the vector field of $\varPhi:=\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$, deriving the direction to follow--the desired velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi\in\mathbb{R}^2$ is the gradient, $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind $w\in\mathbb{R}$.

@@

Suppose uncertainty in the form of $\mathbf{w}_k\in\mathbb{R}^m,v_k\in\mathbb{R}$ accounting for the unknown state and output is absent. The non-linear model in Equation~(\ref{eq:fourier}) can be expressed using an equivalent linear discrete time-invariant state-space model
\begin{equation}\label{eq:state-perf}\begin{cases}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+\mathbf{w}_k\\
  y_k&=C\mathbf{q}_k+v_k
\end{cases},\end{equation}
where $y_k\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption. The state $\mathbf{q}$ mimics the original Fourier series coefficients
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}_k&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    1&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_i=\begin{bmatrix}0 & \frac{i}{\xi} \\ -\frac{i^2}{\xi^2} & 0\end{bmatrix},\\
  C&=\left[\begin{array}{cccccc}
    1 & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}_k\in\mathbb{R}^m$ given $m:=2r+1$, $A\in\mathbb{R}^{m\times m}$ is the state transmission matrix, and $C\in\mathbb{R}^m$ is the output matrix. In matrix $A$, the first value is one, $A_i$ is later on the diagonal, and zero in the remainder.

Suppose at time instant $k$ the mission reached stage $i$ and the control input is $\mathbf{u}_k^a=(\mathbf{c}_k,\mathbf{s}_k)$. The control along with the input matrix
\begin{equation}\label{eq:state-control}\begin{split}  
  \mathbf{u}_k&=\begin{bmatrix}g(\mathbf{s}_k)-g(\mathbf{s}_{k-1}) & \mathbf{c}_k-\mathbf{c}_{k-1}\end{bmatrix}^T,\\
  B&=\left[\begin{array}{cccc}
    1& \omega_{i,1}& \cdots& \omega_{i,\rho}  \\
     &            0&       &  \\
     &             & \ddots&  \\
     &             &       & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{s}_k,g(\mathbf{s}_k)$ are defined in Subsection~\ref{sec:computations-model}, $\mathbf{u}_k\in\mathbb{R}^n$ is the control given $n:=1+\rho$, and $B\in\mathbb{R}^{m\times n}$. Moreover, the first item is one, while the others on the first row are gain factors $\omega_{i}\in\mathbb{R}$, quantifying the contribution of a given adjustment to the instantaneous energy consumption.

@@

The energy evolution analysis necessitates the following realistic assumption.
\begin{assm}[Energy evelution periodicity]\label{assm:periodic} 
Given two time instants $k_1,k_2$ s.t. $k_1>k_2$ and a constant value $n\in\mathbb{Z}_{> 0}$, there exist an arbitrary constant displacement $e\in\mathbb{R}_{\geq 0}$
\begin{equation}
  \abs{y_{k}-y_{k+n}}= e\,\,\,\forall k\in[k_1,k_2].
\end{equation}
\end{assm}

@@

Equation~(\ref{eq:state-control}) accounts for the energy due to the computations. The energy due to the adjustments is merely a linear combination of the gain factor $\omega$ and the adjustment. Nevertheless, the change updates the path which will hence affect the reading from the sensors and adjust the state. For instance, a decrement in the adjustment radius of a circle when the TEE is a circle, adds a negative contribution, thus simulates the lowering of instantaneous energy consumption.

@@

The energy cost of the computations (set of tasks being executed on an embedded board on the UAV) is assessed using \powprof{}, an open-source modeling tool presented in our previous work~\cite{seewald2019coarse}, that measures software configurations empirically and builds a computational energy consumption model. Specifically, the tool builds a linear interpolation, one per each task. It requires the user to implement the mission as a ROS system with one or more ROS nodes changing the computational load by node-specific ROS parameters.

The mission $\mathcal{M}$ contains a set of ordered lists with tasks (recall Definition~\ref{def:mission}). These tasks are simulated by ROS nodes $\Psi(\mathbf{s}_i):=\left(\psi_1(s_{i,1}),\dots,\psi_\sigma(s_{i,\sigma})\right)$, in fact they input the desired and output the actual computations
\begin{equation}\label{eq:qos-def}
  \mathbf{s}_i:=\left\{s_{i,1},\dots ,s_{i,\sigma}\mid \psi_j(s_{i,j})\in\mathbb{S}_{i}\,\forall j\in[\sigma]^+\right\},
\end{equation}
where $s_{i,j}:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $j$-th desired computation at stage $i$, $\mathbf{s}_i\in\mathbb{S}_i\subseteq\mathbb{Z}_{\geq 0}^\sigma$ the union of all QoS sets given $\mathbb{S}_i:=\bigcup_{j\in[\sigma]^+}{\mathbb{S}_{i,j}}$.

@@

Given an initial stage $\mathcal{M}_0$ (TEE $\varphi_0$, tasks $\Psi$, computations $\mathbf{s}_0\in\mathbb{S}_0$, and adaptations $\mathbf{c}_0\in\mathbb{C}_0$), the main purpose of the algorithm is to output a set of control input sequence $\mathbf{u}^a:=\{\mathbf{u}_0^a,\mathbf{u}_1^a,\dots\}$ in a valid mission.

\begin{defn}[Valid mission]\label{def:valid}
  A mission is valid if for every stage $\mathcal{M}_{i-1},\,i\in[l]^+$ there exist a control input $\mathbf{u}_k^{a}$ that produce the next stage $\mathcal{M}_i$
  \begin{equation}\begin{split}
    \mathbf{u}^a_{k}=\{(\mathbf{c}_{k},\mathbf{s}_{k})\mid&(\varphi_i(\mathbf{p}_{k-1},\mathbf{c}_{i-1}),\Psi(\mathbf{s_{i-1}}))\in\mathcal{M}_{i-1}\\
    &\Longrightarrow(\varphi_i(\mathbf{p}_{k},\mathbf{c}_{i}),\Psi(\mathbf{s}_{i}))\in\mathcal{M}_{i}\}.
  \end{split}\end{equation}
\end{defn}

>>>>

\begin{figure*}[t]
  \centering
  \input{\figpath/icra-2021_overview2.tikz}
  \caption{The mission consists of circles and lines covering a polygon. The UAV heads to $\varphi_0$ initially, using the desired velocity vector $\dot{\mathbf{p}}_d$. It later heads to $\varphi_0^-$ reducing the radius $c_{0,1}$ to satisfy the battery constraints (indeed the energy coefficients $\mathbf{q}_{k_2}$ evolution is higher then $\mathbf{q}_{k_3}$). The UAV then converges to $\varphi_1$ in stage $\mathcal{M}_1$, $\varphi_2$ in stage $\mathcal{M}_2$, and so on (the circle $\varphi_2$ is not visible in the figure; it connects $\varphi_1$ and $\varphi_3$).}
  \label{fig:overview}
\end{figure*}

@@

In the spirit of reducing costs, and resources, we showcase the algorithm using the problem of dynamic mission planning for a precision agriculture fixed-wing UAV. Such a scenario is often put into practice~\cite{hajjaj2014review} with ground mobile robots used for harvesting~\cite{qingchun2012study,dong2011development, de2011design, aljanobi2010setup, li2008analysis, edan2000robotic}, and UAVs for preventing damage and ensuring better crop quality~\cite{puri2017agriculture, daponte2019review}. The mission consists of a UAV flying in circles and lines covering a polygon, detecting obstacles using a convolutional neural network (CNN), and informing grounded mobile robots employed for future harvesting--a survey mission optimized for the craft's dynamics. The algorithm plans the mission controlling the processing rate and the radius of the circles, affecting the distance between the lines. Figure~\ref{fig:overview} shows an initial slice of such a mission. The UAV first heads to a circular TEE with a given radius, which is later reduced, as e.g., windy weather required the adjustment of the control input to avoid battery depletion. Computations might also significantly impact the battery. Data indicates a potential extension of up to 13 minutes over an hour by merely switching to the lowest computations.

@@

  The \emph{mission} is the union of all the stages, along with a function $\lambda:\mathbb{R}^2\rightarrow\mathcal{M}_i$ which maps a point $\mathbf{p}\in\mathbb{R}^2$ of a UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ w.r.t. some inertial navigation frame $\mathcal{O}_W$ to a specific stage. For simplicity the system is sampled discrete-time and $l$ is the total number of stages
  \begin{equation}
    \mathcal{M}:=(\lambda(\mathbf{p}),\bigcup_{i\in[l]}{\mathcal{M}_i}),
  \end{equation}
  the algorithm inputs $\mathcal{M}$, initial position, and energy coefficients guess, and outputs the new position, the instantaneous energy consumption, and the control input--an action performed evolving the mission state.
\end{defn}

@@

The set
\begin{equation}\label{eq:area}
  \mathcal{P}_i:=\{\mathbf{p}_k\mid\varphi_i(\mathbf{p}_k,c_{i,1},\dots,c_{i,\rho})\in\mathbb{C}_i\},
\end{equation}
delimits the area where the $i$-th TEE $\varphi_i$ is free to evolve using $\rho$ adjustments $\mathbf{c}_i:=c_{i,1},...,c_{i,\rho}$, being the TEE satisfied for all the approaching points $\varphi_i\rightarrow\mathbb{C}_i$ (gray area in Figure~\ref{fig:overview}).

The algorithm uses the concept to select the adjustments s.t. $\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$ has the highest energy value. It guides the UAV to the new position $\mathbf{p}_{k+1}$ using the vector field of $\varPhi:=\varphi_i(\mathbf{p}_k,\mathbf{c}_i^0)$, deriving the direction to follow--the desired velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}_k):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi$ is the gradient, $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence. The direction the velocity vector $\dot{\mathbf{p}}_d$ is pointing at is generally different from the course heading due to the atmospheric interference, such as wind $w\in\mathbb{R}$.

@@

Suppose uncertainty in the form of $\mathbf{w}_k\in\mathbb{R}^m,v_k\in\mathbb{R}$ accounting for the unknown state and output is absent. The non-linear model in Equation~(\ref{eq:fourier}) can be expressed using an equivalent linear discrete time-invariant state-space model
\begin{equation}\label{eq:state-perf}\begin{cases}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+\mathbf{w}_k\\
  y_k&=C\mathbf{q}_k+v_k
\end{cases},\end{equation}
where $y_k\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption. The state $\mathbf{q}$ mimics the original Fourier series coefficients
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}_k&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    1&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_i=\begin{bmatrix}0 & i/\xi \\ -i^2/\xi^2 & 0\end{bmatrix},\\
  C&=\left[\begin{array}{cccccc}
    1 & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}_k\in\mathbb{R}^m$ given $m:=2r+1$, $A\in\mathbb{R}^{m\times m}$ is the state transmission matrix, and $C\in\mathbb{R}^m$ is the output matrix. In matrix $A$, the first value is one, $A_i$ is later on the diagonal, and zero in the remainder.

Suppose at time instant $k$ the mission reached stage $i$ and the control input is $\mathbf{u}_k^a=(\mathbf{c}_k,\mathbf{s}_k)$. The control along with the input matrix
\begin{equation}\label{eq:state-control}\begin{split}  
  \mathbf{u}_k&=\begin{bmatrix}g(\mathbf{s}_k)-g(\mathbf{s}_{k-1}) & \mathbf{c}_k-\mathbf{c}_{k-1}\end{bmatrix}^T,\\
  B&=\left[\begin{array}{cccc}
    1& \omega_{i,1}& \cdots& \omega_{i,\rho}  \\
     &            0&       &  \\
     &             & \ddots&  \\
     &             &       & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{s}_k,g(\mathbf{s}_k)$ are defined in Subsection~\ref{sec:computations-model}, $\mathbf{u}_k\in\mathbb{R}^n$ is the control given $n:=1+\rho$, and $B\in\mathbb{R}^{m\times n}$. Moreover, the first item of $B$ is one, while the others on the first row are gain factors $\omega_{i}\in\mathbb{R}$, quantifying the contribution of a given adjustment to the instantaneous energy consumption.

@@

Equation~(\ref{eq:state-control}) accounts for the energy due to the computations. The energy due to the adjustments is merely a linear combination of the gain factor $\omega$ and the adjustment. Nevertheless, the change updates the path which will hence affect the reading from the sensors and adjust the state. For instance, a decrement in the adjustment radius of the circle when the TEE is a circle, adds a negative contribution, thus simulates the lowering of instantaneous energy consumption.

The energy consumption modeling of the mission necessitates the following realistic assumption.
\begin{assm}[Energy evelution periodicity]\label{assm:periodic} 
Given two time instants $k_1,k_2$ s.t. $k_1>k_2$ and a constant value $n\in\mathbb{Z}_{> 0}$, there exist an arbitrary constant displacement $e\in\mathbb{R}_{\geq 0}$
\begin{equation}
  \abs{y_{k}-y_{k+n}}= e\,\,\,\forall k\in[k_1,k_2].
\end{equation}
\end{assm}

@@

@@

The energy cost of the computations (set of tasks being executed on an embedded board on the UAV) is assessed using \powprof{}, an open-source modeling tool presented in our previous work~\cite{seewald2019coarse}. The tool measures software configurations empirically and builds a computational energy consumption model: a linear interpolation, one per each task. It requires the user to implement the mission as a ROS system with one or more ROS nodes changing the computational load by node-specific ROS parameters.

Specifically, the mission $\mathcal{M}$ contains a set of ordered lists with tasks (recall Definition~\ref{def:mission}). These tasks are simulated by ROS nodes $\Psi(\mathbf{s}_i):=\left(\psi_1(s_{i,1}),\dots,\psi_\sigma(s_{i,\sigma})\right)$, in fact they input the desired and output the actual computations
\begin{equation}\label{eq:qos-def}
  \mathbf{s}_i:=\left\{s_{i,1},\dots ,s_{i,\sigma}\mid \psi_j(s_{i,j})\in\mathbb{S}_{i}\,\forall j\in[\sigma]^+\right\},
\end{equation}
where $s_{i,j}:\mathbb{Z}_{\geq 0}\rightarrow\mathbb{Z}_{\geq 0}$ returns the $j$-th desired computation at stage $i$, $\mathbf{s}_i\in\mathbb{S}_i\subseteq\mathbb{Z}_{\geq 0}^\sigma$ the union of all QoS sets given $\mathbb{S}_i:=\bigcup_{j\in[\sigma]^+}{\mathbb{S}_{i,j}}$.

@@

Given an initial stage $\mathcal{M}_0$ (TEE $\varphi_0$, tasks $\Psi$, computations $\mathbf{s}_0\in\mathbb{S}_0$, and adaptations $\mathbf{c}_0\in\mathbb{C}_0$), the main purpose of the algorithm is to output a control input sequence $\mathbf{u}^a:=\{\mathbf{u}_0^a,\mathbf{u}_1^a,\dots\}$ in a valid mission.

\begin{defn}[Valid mission]\label{def:valid}
  A mission is valid if for every stage $\mathcal{M}_{i-1},\,i\in[l]^+$ there exist a control input $\mathbf{u}_k^{a}$ that produce the next stage $\mathcal{M}_i$
  \begin{equation}\begin{split}
    \mathbf{u}^a_{k}=\{&(\mathbf{c}_{k},\mathbf{s}_{k})\mid\,\exists n\in\mathbb{Z}_{>0},\\&(\varphi_{i-1}(\mathbf{p}_{k-n},\mathbf{c}_{k-n}),\Psi(\mathbf{s}_{k-n}))\in\mathcal{M}_{i-1}\\
    &\Longrightarrow(\varphi_i(\mathbf{p}_{k},\mathbf{c}_{k}),\Psi(\mathbf{s}_{k}))\in\mathcal{M}_{i}\}.
  \end{split}\end{equation}
\end{defn}
